Combinatorics of the geometry of Wilson loop diagrams I: equivalence classes via matroids and polytopes

### 2019-07-04 GWLDCI.tex
Proofread intro

This paper is the first in a two part series investigating the combinatorics and geometry underlying SYM $N=4$ theory. This series lays out several results about the relationship between the Wilson loop diagrams and the positroid cells that give a CW complex structure to the positive Grassmanian. This paper expands on an observation from \cite{wilsonloop}, that shows that certain families of Wilson loop diagrams correspond to the same positroid cell. This paper gives a way of enumerating the number of different Wilson loop diagrams that correspond to a given positroid cell. Furthermore, it shows that the number of cells in the image of the map to positive Grassmannians is given by counting the directions in associahedra.

@@

Meanwhile, there has been a different body of work studying SYM $N=4$ theory from the point of view of Feynman integrals in twistor space \cite{Adamo:2011pv, Britto:2005fq, Cachazo:2004kj}. These integrals are calculated via Wilson loops. The first author and Amat uncovered a connection between the Feynman integrals developed in this literature and positroids, defined  as matroids that are realized as an element of $\Gr(k,n)$ \cite{wilsonloop}. With the second author and others, this work was extended to study these Feynman integrals in terms of positroid cells and the positive Grassmanians \cite{casestudy, non-orientable}. Others have tried different approaches to define a geometric space associated to these integrals in a manner similar to the Amplituhedron \cite{Amplituhedronsquared, HeslopStewart}.

@@

In this paper, we focus on the correspondence between Wilson loop diagrams and positroids, first defined in \cite{wilsonloop}. In that paper, one of us with Amat show that there are certain families of diagrams, specifically those with exact subdiagrams, that do map to the same positroid cell. In \cite{casestudy} some of us explictly show that not every positroid cell of a given dimension corresponds to a Wilson loop diagram. In this paper, we better enumerate this correspondence. We show that the map from Wilson loop diagrams to postroid cells is one to one exactly on the subset of diagrams with no non-trivial exact subdiagrams. We show that exact diagrams correspond to uniform matroids, and that contraction by exact subdiagrams has special matroidal properties. Further, given a Wilson loop diagram with a non-trivial exact subdiagram, we count the total number of Wilson loop diagrams that correspond to that same positroid cell. While we do not count the number of distint positroid cells that correspond to Wilson loop diagrams, we do show that this enumeration is equivalent to enumerating the number of directions of an associahedron.

@@

We proceed in Section~\ref{section overall background} with the required definitions for Wilson loop diagrams (subsection~\ref{section WLD background}) and matroids (subsection~\ref{sec matroid background}).  Wilson loop diagrams  are said to be admissible when they correspond to $N^kMHV$ diagrams, or mathematically according to Definition~\ref{admisdfn}.  With this, the key definition is Definition~\ref{equivdfn} giving the equivalence relation on admissible Wilson loop diagrams.  The connection to matroids is stated precisely in Theorem \ref{thm WLD defines matroid}.

@@

Having identified the equivalence classes and their sizes, the next natural question is to identify the number of inequivalent diagrams under this equivalence relation.  The main result of Section~\ref{sec associahedron} is to show that the set of admissible Wilson loop diagrams (up to equivalence) is in bijection with the set of faces in the associahedron (up to paralellness).  Towards this result we go over some relevant background on polytopes in subsection~\ref{sec polytope background}, and then give the main result in subsection~\ref{sec associahedron results} in the form of two propositions: Propositions~\ref{prop easy way} and~\ref{prop hard way}.

@@

@@

Propagators are undirected, so $p = (i,j) = (j,i)$. In order to fix a convention, we write $p = (i,j)$ with  $i +1 \leq j$ relative to the first vertex.

>>>>

This paper is the first in a two part series investigating the combinatorics and geometry underlying SYM $N=4$ theory. This series lays out several results about the relationship between Wilson loop diagrams and the positroid cells that give a CW complex structure to the positive Grassmanian. This paper expands on an observation from \cite{wilsonloop}, that shows that certain families of Wilson loop diagrams correspond to the same positroid cell. This paper gives a way of enumerating the number of different Wilson loop diagrams that correspond to a given positroid cell. Furthermore, it shows that the number of cells in the image of the map to positive Grassmannians is given by counting the directions \hlfix{in}{of a?} associahedra.

@@

Meanwhile, there has been a different body of work studying SYM $N=4$ theory from the point of view of Feynman integrals in twistor space \cite{Adamo:2011pv, Britto:2005fq, Cachazo:2004kj}. These integrals are calculated via Wilson loops. The \hlfix{first author and Amat}{At other points, we say "some of us" instead of specifying like this. Choose a convention} uncovered a connection between the Feynman integrals developed in this literature and positroids, defined  as matroids that are realized as an element of $\Gr(k,n)$ \cite{wilsonloop}. With the second author and others, this work was extended to study these Feynman integrals in terms of positroid cells and the positive Grassmanians \cite{casestudy, non-orientable}. Others have tried different approaches to define a geometric space associated to these integrals in a manner similar to the Amplituhedron \cite{Amplituhedronsquared, HeslopStewart}.

@@

In this paper, we focus on the correspondence between Wilson loop diagrams and positroids, first defined in \cite{wilsonloop}. In that paper, \hlfix{one of us}{see previous comment about self reference convention} with Amat show that there are certain families of diagrams, specifically those with exact subdiagrams, that do map to the same positroid cell. In \cite{casestudy} some of us explictly show that not every positroid cell of a given dimension corresponds to a Wilson loop diagram. In this paper, we better enumerate this correspondence. We show that the map from Wilson loop diagrams to postroid cells is one to one exactly on the subset of diagrams with no non-trivial exact subdiagrams. We show that exact diagrams correspond to uniform matroids, and that contraction by exact subdiagrams has special matroidal properties. Further, given a Wilson loop diagram with a non-trivial exact subdiagram, we count the total number of Wilson loop diagrams that correspond to that same positroid cell. While we do not count the number of distinct positroid cells that correspond to Wilson loop diagrams, we do show that this enumeration is equivalent to enumerating the number of directions of an associahedron.

@@

We proceed in Section~\ref{section overall background} with the required definitions for Wilson loop diagrams (subsection~\ref{section WLD background}) and matroids (subsection~\ref{sec matroid background}).  Wilson loop diagrams  are said to be admissible when they correspond to $N^kMHV$ diagrams, or mathematically according to Definition~\ref{admisdfn}. The other key definition in this section is Definition~\ref{equivdfn}, which gives an equivalence relation on admissible Wilson loop diagrams.  The connection to matroids is stated precisely in Theorem \ref{thm WLD defines matroid}.

@@

Having identified the equivalence classes and their sizes, the next natural question is to identify the number of inequivalent diagrams under this equivalence relation.  The main result of Section~\ref{sec associahedron} is Theorem \ref{thm:count inequiv diagrams} which shows that the set of admissible Wilson loop diagrams (up to equivalence) is in bijection with the set of faces in the associahedron (up to paralellness).  Towards this result we go over some relevant background on polytopes in subsection~\ref{sec polytope background}, and then give the main result in subsection~\ref{sec associahedron results} in the form of two propositions: Propositions~\ref{prop easy way} and~\ref{prop hard way}.

@@

This section gives the relevant physical and combinoatorial background for this paper. Section \ref{section WLD background} defines Wilson loop diagrams as combinatorial objects, and presents relevant definitions and results from previous work. Section \ref{sec matroid background} presents some background on matroids and some known result relating Wilson loop diagrams to matroids.

@@

Propagators are undirected, so $p = (i,j) = (j,i)$. In previous work, we have used the convention, we write $p = (i,j)$ with  $i +1 \leq j$ relative to the first vertex. \sanote{Susama stopped here}
